---
title: "NYC_squirrels"
author: "Alina Sansevich"
date: "April 10, 2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# New York City Squirrels Census

I got the idea for this analysis from: https://www.springboard.com/blog/15-fun-datasets-to-analyze/

@@@@@ TO-DO:
WRITE ABOUT THE SQUIRRELS CENSUS = WHERE DOES THIS DATASET COME FROM
https://www.thesquirrelcensus.com/

This is my first data analysis project with R, and also my first R Markdown document.

```{r}
library(readr)
library(dplyr)
library(lubridate) # to use function month
library(ggplot2)
library(ggthemes)
```

Let's start by loading the dataset, and creating the dataframe nyc_squirrels:

```{r}
# 1) load the dataset
nyc_squirrels <- read.csv('nyc_squirrels.csv')
```

## Data Wrangling: 
```{r}
# 2) check str, stats, NAs, data types...
str(nyc_squirrels)
```

As `str(nyc_squirrels)` shows @@@@@, this dataset contains data that could be classified as belonging to 1 of 4 different categories:

* location (_where were the squirrels sighted/seen?_), 
* features (_what did the squirrels look like?_, _how old were they?_),
* behavior (_what were they doing?_),
* time (_when were they sighted/seen?_).

I will organize my analysis around these 4 categories of information. 

The date column's data type is Integer, I will change it to 'Date'.
```{r}
# change nyc_squirrels$date data type to character
nyc_squirrels$date <- as.character(nyc_squirrels$date)

# define function add_slash: takes the date in format: "10142018" (class Character)
# and returns it in format: "10/14/2018" (class Character) 
add_slash <- function(x){
  day <- substr(x, 3, 4)
  month <- substr(x, 1, 2)
  year <- substr(x, 5, 8)
  return <- paste(month, day, year, sep="/")
}
# change each date value with apply add_slash
new_date <- apply(nyc_squirrels[,6,drop=F], 1, add_slash)
nyc_squirrels$date <- new_date
# finally, change data type from 'Character' to 'Date'
nyc_squirrels$date <- as.Date(nyc_squirrels$date, "%m/%d/%Y")
```

In the age column there are "?" values and NAs. This column is a factor, so "?" is a level,
this means that `nyc_squirrels[nyc_squirrels$age == "?" ] <- NA` doesn't change the factor's level. So, to change the factor's level from "?" to NA:

```{r}
levels(nyc_squirrels$age)[levels(nyc_squirrels$age) == '?'] <- NA
summary(nyc_squirrels$age)
```





